### 2016-09-22 article.tex
Jetzt aber, erste Version ist auch gepusht.

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
\usepackage[]{amsmath}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Daniel Kaschek\\University of Freiburg \And
	Marcus Rosenblatt\\University of Freiburg \AND
	Wolfgang Mader\\University of Freiburg \And
	Mirjam Fehling-Kaschek\\University of Freiburg}

\title{Dynamic Modeling and Parameter Estimation in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Daniel Kaschek, Marcus Rosenblatt, Wolfgang Mader, Mirjam Fehling-Kaschek} %% comma-separated
\Plaintitle{Dynamic Modeling and Parameter Estimation in R} %% without formatting
\Shorttitle{Dynamic Modeling} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{Dynamic models, parameter estimation, code generation, maximum likelihood}
\Plainkeywords{Dynamic models, parameter estimation, code generation, maximum likelihood} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Daniel Kaschek\\
  Institute of Physics, University of Freiburg\\
  Hermann-Herder-Str.~3\\
  79104 Freiburg, Germany\\
  E-mail: \email{daniel.kaschek@physik.uni-freiburg.de}
  %URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{} commands.

\section{Theoretical background}

\subsection{Ordinary differential equation models}

\subsection{Maximum-likelihood method and non-linear optimization}

\subsection{Uncertainty analysis}

\section{Implementation and design principles}

\subsection{Model formulation}
\subsection{Observation functions}
\subsection{Parameter transformations}
\subsection{Multi-conditional parameter estimation}

\section{Three-compartment model of bile acid transport}

The following model is a simplified version of the dynamic model on bile acid transport published in [REF]. Bile acids are produced in the liver. They are necessary for the digestion of fat and oil. In the liver, bile acids are taken up in hepatocytes (liver cells) by specific transporter molecules. Clearance occurs either via canalicular or sinusoidal export, i.e.~export to the bile transportation system of bile canaliculi or the intercellular space.

To study bile clearance, experiments with hepatocytes or hepatocyte-like cell lines are performed \textit{in-vitro}. Cells in a Petri dish stick together to a monolayer of cells, forming bile canaliculi in between cells. They adhere to the plate and are surrounded by a buffer providing the cells with nutrients. To measure bile acids, here \textit{taurocholic acid} (TCA), radioactively labeled TCA is used.

For the mathematical description of bile flow, a three-compartment differential equation model is used. TCA is pipetted into the buffer compartment (\code{TCA_buffer}) from where it is transported into the cells (\code{TCA_cell}). Intracellular TCA is exported back to the buffer or the canalicular compartment (\code{TCA_cana}). Finally, canalicular TCA flows back into the buffer compartment. These processes give rise to the following differential equations:
\begin{equation}
	\begin{aligned}
		\frac{\rm d}{ {\rm d}t} \verb!TCA_buffer! &= \verb!-import * TCA_buffer + export_sinus * TCA_cell!\\
		& \quad \verb! + reflux * TCA_cana!\\
		\frac{\rm d}{ {\rm d}t} \verb!TCA_cana! &= \verb!export_cana * TCA_cell - reflux * TCA_cana!\\
		\frac{\rm d}{ {\rm d}t} \verb!TCA_cell! &= \verb!import * TCA_buffer - export_sinus * TCA_cell!\\
		&\quad \verb! - export_cana * TCA_cell!.
	\end{aligned}
	\label{eq:ode}
\end{equation}

\subsection{Simulation and prediction}

Each transportation process is modeled by mass-action kinetics. The implementation in \pkg{dMod} would be as follows:
\begin{CodeChunk}
	\begin{CodeInput}
# Add reactions
reactions <- NULL
reactions <- addReaction(reactions, "TCA_buffer", "TCA_cell",
			 rate = "import*TCA_buffer",
			 description = "Uptake")
reactions <- addReaction(reactions, "TCA_cell", "TCA_buffer",
			 rate = "export_sinus*TCA_cell",
			 description = "Sinusoidal export")
reactions <- addReaction(reactions, "TCA_cell", "TCA_cana",
			 rate = "export_cana*TCA_cell",
			 description = "Canalicular export")
reactions <- addReaction(reactions, "TCA_cana", "TCA_buffer",
			 rate = "reflux*TCA_cana",
			 description = "Reflux into the buffer")

# Translate into ODE model
mymodel <- odemodel(reactions, modelname = "bamodel")

# Generate prediction function from ODE model
x <- Xs(mymodel, condition = NULL)
	\end{CodeInput}
\end{CodeChunk}
The reactions are collected in an \code{eqnlist} object. The \code{odemodel()} command composes the single reactions to an ODE system and auto-generates the \proglang{C} code which is used by the \pkg{deSolve} package to evaluate the ODE. Prediction functions generated by the \code{Xs()} command are wrapper functions for the call of \code{ode()}. The "s" in \code{Xs()} refers to "sensitivities", i.e the prediction function not only computes and returns the model prediction but sensitivities, too. The usage of the prediction function is illustrated by the following code chunk. Time points are defined between 0 and 50, numeric values are assigned to all model parameters. Finally the prediction function is called and the result, a prediction list, is plotted, shown in Figure~\ref{fig:prediction}A.

\begin{CodeChunk}
\begin{CodeInput}
times <- seq(0, 50, .1)
pars <- c(TCA_buffer = 1,
          TCA_cell = 0,
	  TCA_cana = 0,
	  import = 0.2,
	  export_sinus = 0.2,
	  export_cana = 0.04,
	  reflux = 0.1)

# Generate the prediction
out <- x(times, pars)
plot(out)

# Get sensitivities for the dynamic parameters
out <- getDerivs(x(times, pars[4:7], fixed = pars[-(4:7)]))
plot(out)
\end{CodeInput}
\end{CodeChunk}

\begin{figure}[ht]
	\includegraphics[width = \textwidth]{images/figure1}
	\caption{Output of the prediction function. \textbf{(A)} Prediction of the TCA states. \textbf{(B)} Sensitivities of the three TCA states with respect to the rate parameters.}
	\label{fig:prediction}
\end{figure}

Figure~\ref{fig:prediction}A shows the uptake of TCA in the cell and canaliculi, saturating around $t = 50$. The prediction parametrically depends on initial values and rate parameters. Figure~\ref{fig:prediction}B shows the model sensitivities $\frac{\partial x}{\partial p}$ for the rate parameters.

\subsection{Observation function and simulated data}
The three dynamic states, \code{TCA_buffer}, \code{TCA_cell} and \code{TCA_cana} cannot be directly measured. In practice, the cellular fraction, i.e. both cells and canaliculi, can be separated from the buffer. The radioactivity in the two fractions is determined independently. In summary, we find the following relation between the radioactive counts and the dynamic states of our ODE model:
\begin{equation}
	\begin{aligned}
		\verb!buffer! &= \verb!s * TCA_buffer!\\
		\verb!cellular! &= \verb!s * (TCA_cana + TCA_cell)!
	\end{aligned}
	\label{}
\end{equation}
The scaling factor \code{s} translates amounts into radioactive counts. The fact that the cellular fraction contains cells as well as canaliculi is represented by the sum. The implementation in \pkg{dMod} follows.

First, observables are defined by an \code{eqnvec} object from which an observation function \code{g} is generated by the \code{Y()} command. The \code{Y()} command needs to be informed which of the symbols are variables (dynamic states) or parameters. For convenience, this can be done passing the \code{reactions} that underly our ODE model.
\begin{CodeChunk}
\begin{CodeInput}
# Generate observation function
observables <- eqnvec(
  buffer = "s*TCA_buffer",
  cellular = "s*(TCA_cana + TCA_cell)"
)

g <- Y(observables, reactions, condition = NULL,
       compile = TRUE, modelname = "obsfn")

\end{CodeInput}
\end{CodeChunk}
The data we are going to simulate is for an efflux experiment, i.e. time point $t = 0$ reflects the distribution of TCA after uptake. Hence, we reset some of the parameter values. Initial values are set to steady state values. The incubation buffer is replaced by fresh buffer, meaning that \code{TCA_buffer} is set to 0. Finally, we assume that \code{s = 1000}. The prediction of the observables is made by concatenation of \code{g} and \code{x} via the concatenation operator \code{"*"} being defined in \pkg{dMod} for prediction, observation and parameter functions.
\begin{CodeChunk}
\begin{CodeInput}
# Reset parameter values
pars["TCA_cell"] <- 0.3846154
pars["TCA_cana"] <- 0.1538462
pars["TCA_buffer"] <- 0
pars["s"] <- 1e3

out <- (g*x)(times, pars, conditions = "standard")
\end{CodeInput}
\end{CodeChunk}
Finally, we obtain the simulated data by taking a subset of the prediction. At this point we make use of the "conditions" argument which allows to request a prediction for a vector of conditions, if provided by the prediction function\footnote{Both, \code{x} and \code{g} have been initialized with \code{condition = NULL} which means they are generic prediction functions and yield the same prediction for any of the requested conditions. The concept is explained in Section~\ref{sec:conditions}.}. Data uncertainties $\sigma$ are derived by the Poisson nature of radioactive count experiments, i.e. $\sigma_x = \sqrt{x}$. To avoid division by 0, the minimal $\sigma$-value is set to 1. Random values are added to the predicted values to simulate observation noise. In the end, the \code{data.frame} is converted into a \code{datalist} object.
\begin{CodeChunk}
\begin{CodeInput}
# Simuate data
timesD <- c(0.1, 1, 3, 7, 11, 15, 20, 41)
datasheet <- subset(wide2long(out),
                    time %in% timesD & name %in% names(observables))
datasheet$sigma <- sqrt(datasheet$value + 1)
datasheet$value <- rnorm(nrow(datasheet), datasheet$value, datasheet$sigma)

data <- as.datalist(datasheet)

plot(out, data)
\end{CodeInput}
\end{CodeChunk}
Both, the simulated data and the model prediction from which the data is derived are shown in Figure~\ref{fig:observation}.
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure2}
	\caption{Model prediction of the internal states and observables. Simulated data is shown as dots with error bars.}
	\label{fig:observation}
\end{figure}

\subsection{Parameter transformation}
Parameter transformations play a crucial role in the set-up of \pkg{dMod}. They can have several purposes such as fixing paramer values, implementing parameter bounds, including steady-state constraints or mapping parameters to different conditions.

First, we use parameter transformations to constrain all parameters to be positive or zero because all our parameters are either amounts or rate parameters. The parameter transformation is generated by the \code{P()} command from an \code{eqnvec} object, explicitly stating the relation between the \textbf{inner} parameters, i.e.~the parameter values that are evaluated within the model, and the \textbf{outer} parameters, i.e.~the parameter values handed by the user or by an optimizer, from which the inner parameters are computed. The corresponding code reads:
\begin{CodeChunk}
\begin{CodeInput}
p <- P(
  trafo = eqnvec(
    TCA_buffer = "0",
    TCA_cell = "exp(TCA_cell)",
    TCA_cana = "exp(TCA_cana)",
    import = "exp(import)",
    export_sinus = "exp(export_sinus)",
    export_cana = "exp(export_cana)",
    reflux = "exp(reflux)",
    s = "exp(s)"
  ),
  condition = "standard"
)

outerpars <- getParameters(p)
pouter <- structure(rep(-1, length(outerpars)), names = outerpars)
plot((g*x*p)(times, pouter), data)
\end{CodeInput}
\end{CodeChunk}
The vector \code{outerpars} is the collection of all symbols on the right-hand side of \code{trafo}. It agrees with \code{names(pars)} except for \code{TCA_buffer}, which is fixed to 0 by the transformation. However, the meaning of the parameters has changed since now their values are on a log-scale. All three functions, the observation function, prediction function and parameter transformation can be concatenated to one new prediction function, \code{g*x*p} which takes times and values of the outer parameters to predict internal and observable states. The model prediction generated by \code{pouter} is shown in Figure~\ref{fig:gxp}.
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure3}
	\caption{Prediction of internal and observed states. All values of the outer parameters have been set to $-1$. Simulated data points are shown as dots with error bars.}
	\label{fig:gxp}
\end{figure}

\subsection{Objective function and model fitting}
For nomally distributed measurement noise, maximum-likelihood estimation is equivalent to least-squares estimation. A least-squares objective function can be generated by the \code{normL2()} command which requires a \code{datalist} object, in our case \code{data}, and a prediction function, in our case \code{g*x*p}.

Frequently, non-identifiable parameters are encountered in non-linear dynamic systems. To prevent the optimizer from taking extreme parameter values for which the ODE solver crashes, a general $L_2$ prior can be used on all parameters, being implemented by the \code{constraintL2()} command. The command returns an \code{objfn} objection, as does \code{normL2()}. Both can be added by the \code{"+"} operator.

The following code illustrates the implementation of the objective function and how it is used with the \code{trust()} optimizer from the \pkg{trust} package to obtain a model fit, shown in Figure~\ref{fig:myfit}.

\begin{CodeChunk}
\begin{CodeInput}
obj <- normL2(data, g*x*p) + constraintL2(pouter, sigma = 10)
myfit <- trust(obj, pouter, rinit = 1, rmax = 10)

plot((g*x*p)(times, myfit$argument), data)
\end{CodeInput}
\end{CodeChunk}

\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure4}
	\caption{Prediction of internal and observed states after optimization of the objective function. Simulated data points are shown as dots with error bars.}
	\label{fig:myfit}
\end{figure}

Besides non-identifiability of parameters, local optima is another feature of non-linear optimization. The \code{trust()} optimizer employs derivative information and therefore, if starting within a certain radius around a local optimum, is very efficient in finding it back. Once an optimum is found, we can be confident that there is no deeper point around. However, to be confident that an optimum is the globally best solution, we might want to scatter starting points for optimization runs all over the parameter space. The \pkg{dMod} package provides the \code{mstrust()} function based on \code{trust()} to do a multi-start search:
\begin{CodeChunk}
\begin{CodeInput}
out_mstrust <- mstrust(obj, pouter, rinit = 1, rmax = 10, iterlim = 500,
                       sd = 4,
                       cores = 4, fits = 50)

myframe <- as.parframe(out_mstrust)

plotValues(myframe, value < 100)
plotPars(myframe, value < 100)
\end{CodeInput}
\end{CodeChunk}
Here, we have searched according to $\vec p_i = \vec p_0 + \Delta\vec p_i$ where $\vec p_0$ is the center, in our case \code{pouter}, $\Delta\vec p_i\sim N(0, \sigma^2)$ is a random parameter vector taken from a normal distribution, in our case $\sigma = 4$, and the index $i$ runs from 1 to \code{fits = 50}. The \code{mclapply()} command from the \pkg{parallel} package is used internally to run fits in parallel, here \code{cores = 4}. The result of \code{mstrust()} is a list of all returned values of \code{trust()}. To extract the final objective value, parameter values, convergence information and the number of interations, \code{as.parframe()} is used. The multi-start approach identifies four local optima, see Figure~\ref{fig:mstrust}, which yield almost the same objective value, Figure~\ref{fig:mstrust}A.
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure5}
	\caption{Result of multi-start fitting procedure. \textbf{(A)} Fits have been sorted by increasing least squares value. Four optima were found with almost identical objective value. \textbf{(B)} The parameter values for different optima are shown in different colors }
	\label{fig:mstrust}
\end{figure}
Despite the similar objective value, the optima are not close to each other in parameter space, as being illustrated by Figure~\ref{fig:mstrust}B. From the initial value parameters \code{TCA_cell} and \code{TCA_cana} it seems that they are exchangable. Where \code{TCA_cell} is high, \code{TCA_cana} is low and vice versa. This is not surprising because the observed cellular TCA amount is the sum of both. A new experiment needs to be designed to distinguish one situation from another.

\subsection{Working with several conditions}\label{sec:conditions}

In practice, the canaliculi only form a closed compartment if Ca$^{2+}$/Mg$^{2+}$ ions are present in the buffer. Therefore, if the experiment is repeated with Ca$^{2+}$/Mg$^{2+}$-free efflux buffer, the contents of the canaliculi escapes quickly into the buffer compartment and the observables \code{buffer} and \code{cellular} effectivly reflect \code{TCA_buffer + TCA_cana} and \code{TCA_cell}. Mathematically, two experimental conditions which differ only by the \code{reflux} parameter need to be combined in one objective function.

Like before, we generate a simulated data set. Then, a parameter transformation for the additional condition is set up and the parameter space is explored by a multi-start fit.

\begin{CodeChunk}
\begin{CodeInput}
pars["reflux"] <- 1e3
out <- (g*x)(times, pars, conditions = "open")

datasheet <- subset(wide2long(out),
		    time %in% timesD & name %in% names(observables))
datasheet$sigma <- sqrt(datasheet$value + 1)
datasheet$value <- rnorm(nrow(datasheet), datasheet$value, datasheet$sigma)

data <- data + as.datalist(datasheet)
\end{CodeInput}
\end{CodeChunk}
Similar to objective functions and observation/prediction/parameter functions, as we will see, datalists can be combined by the \code{"+"} operator.

To add a condition to the parameter transformation function, we use the equations of the standard condition as a template for the "open" condition. Parameter transformation functions for different conditions are combined by the \code{"+"} operator.

\begin{CodeChunk}
\begin{CodeInput}
trafo <- summary(p)$standard$equations
trafo["reflux"] <- "exp(reflux_open)"
p <- p + P(trafo, condition = "open")
\end{CodeInput}
\end{CodeChunk}

Compared to the situation before we have gained the outer parameter \code{reflux_open}. Both transformations, "standard" and "open", evaluate the same vector of outer parameters meaning that they yield the same values for all but \code{reflux} and \code{reflux_open}. The prediction function \code{g*x} is generic in the sense that \code{condition = NULL} wheras the concatenation \code{g*x*p} has the conditions "standard" and "open", evaluating the identical function \code{g*x} on two parameter vectors.

We define un updated objective function:
\begin{CodeChunk}
\begin{CodeInput}
outerpars <- getParameters(p)
pouter <- structure(rep(-1, length(outerpars)), names = outerpars)

obj <- normL2(data, g*x*p) + constraintL2(pouter, sigma = 10)
\end{CodeInput}
\end{CodeChunk}

Then we start 50 fits around \code{pouter}. The list of fits is simplified to a \code{parframe} and by the \code{as.parvec()} function, the parameter vector (of the best fit) is extracted from the \code{parframe}. The best fit is used to make a prediction and plot it together with the simulated data. All results are shown in Figure~\ref{fig:twoconditions}.

\begin{CodeChunk}
\begin{CodeInput}
out_mstrust <- mstrust(obj, pouter, rinit = 1, rmax = 10, iterlim = 500,
		       sd = 4,
		       cores = 4, fits = 50)
myframe <- as.parframe(out_mstrust)
plotValues(myframe, value < 100)
plotPars(myframe, value < 100)

bestfit <- as.parvec(myframe)
plot((g*x*p)(times, bestfit), data)
\end{CodeInput}
\end{CodeChunk}
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure6}
	\caption{Result of multi-start fitting procedure with two experimental conditions. \textbf{(A)} Model prediction of the best fit and the simulated data are shown in different colors. \textbf{(B)} Fits have been sorted by increasing least squares value. The lowest value clearly separates from the second plateau. \textbf{(C)} Plotting the parameter values for each of the fits reveals that the second plateau consists of two optima. The lowest plateau however corresponds to a unique optimum.}
	\label{fig:twoconditions}
\end{figure}

Interestingly, with the new experiment the best optimum becomes unique. The log-likelihood difference, i.e.~half the difference between the objective values, is more than 20 between the lowest and the second plateau, which is highly significant. The uniqueness of the lowest plateau is confirmed by Figure~\ref{fig:twoconditions}C which shows no scattering of the black circles.

\subsection{Parameter uncertainty and identifiability}

One might wonder why the optimum is unique as for any choice of the scaling parameter \code{s} we find appropriate values of the TCA initial value parameters that give rise to exactly the same prediction of the observables. The reason for the uniqueness is the parameter $L_2$-constraint that we have added to the objective function. Nonetheless, we will see, that the non-identifiability is still visible in the profile likelihood.

The profile likelihood is computed by the \code{profile()} command. There are several options to control the step-size and accuracy. For convenience the \code{method} option can be used to select between the presets \code{"integrate"} and \code{"optimize"}.

The code
\begin{CodeChunk}
\begin{CodeInput}
profiles <- profile(obj, bestfit, names(bestfit), limits = c(-5, 5), cores = 4)
plotProfile(profiles)
plotPaths(profiles, whichPar = "s")
\end{CodeInput}
\end{CodeChunk}
gives us the result shown in Figure~\ref{fig:pl}.
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure7}
	\caption{Profile likelihood. \textbf{(A)} Profiles of all parameters. Data- and prior contribution to the total objective value are distinguished by line-type. \textbf{(B)} Parameter paths for the scaling parameter \code{s}.}
	\label{fig:pl}
\end{figure}
Computing the profile likelihood, the sum of data contribution, \code{normL2}, and prior contribution, \code{constraintL2}, are optimized under the constraint of a given parameter value for the profiled parameter. In the optimum, data and prior contribution are evaluated separately giving rise to the dashed and dotted lines in Figure~\ref{fig:pl}A. As we had expected, the data contribution to the initial value parameters \code{TCA_cell}, \code{TCA_cana} and the scaling parameter \code{s} is constantly zero. The parameters are structurally non-identifiable.

Each profile corresponds to a certain path in parameter space. The path for the profile of the non-identifiable scaling parameter \code{s} is shown in Figure~\ref{fig:pl}B. It shows a clear coupling of scaling and the initial value parameters: both initial value parameters have to be decreased by the same extent as the scaling parameter is increased to keep the prediction unchanged.

The parameters \code{export_sinus} and \code{reflux_open} exceed the 95\% confidence threshold only to one side. Given the data, the \code{export_sinus} parameter could equally be $-\infty$ (corresponding to an export rate of 0) without changing the likelihood significantly to the worse. A similar statement holds for the \code{reflux_open} parameter which could equally be $\infty$ meaning that we could assume instantaneous draining of the canaliculi for the \code{"open"} condition. The two parameters are practically non-identifiable and allow to reduce the model.

Finally, the parameters \code{import}, \code{export_cana} and \code{reflux} exceed the 95\% confidence threshold in both directions meaning that the parameters have finite confidence intervals. However, the confidence intervals are rather large and we might ask if there is further information that we could use to improve parameter identifiability without generating new data.

\subsection{Steady-state constraints and implicit transformations}

So far we have estimated both initial concentrations, \code{TCA_cell} and \code{TCA_cana}, independently. However, we know that the efflux experiment was just started after completion of the uptake process. Our system runs into a steady state, the buffer is exchanged and the measurement begins. Hence, we can use the steady-state condition as an additional information for the modeling process.

The steady-state relation between \code{TCA_cana} and \code{TCA_cell} can be derived analytically from eq.~\eqref{eq:ode}. It reads
\begin{equation}
	\verb!TCA_cana! = \verb!export_cana * TCA_cell / reflux!
	\label{}
\end{equation}
This relation can be explicitly used in a parameter transformation to express \code{TCA_cana} in terms of other parameters. The dimension of the parameter space is thereby reduced by one. The following implementation shows how we would use the existing transformation function \code{p} to generate an alternative transformation function \code{pSS} to replace \code{p} in the prediction function \code{g*x*p}.
\begin{CodeChunk}
\begin{CodeInput}
pSS <- NULL
trafos <- summary(p)
conditions <- names(trafos)
for (n in conditions) {

  equations <- trafos[[n]]$equations
  equations["TCA_cana"] <- "exp(export_cana)*exp(TCA_cell)/exp(reflux)"
  pSS <- pSS + P(equations, condition = n)

}
\end{CodeInput}
\end{CodeChunk}
We get all the information about the transformations from the \code{summary()} command. The equation for \code{TCA_cana} is substituted by our steady-state constraint. By the \code{"+"} operator, a new parameter transformation function \code{pSS} for our multiple conditions is constructed step-by-step.

Next, we want to implement the steady-state constraint by an implicit parameter transformation. Let $\dot x = f(x, p)$ be our dynamic system. Then, under certain conditions, we find a function $g(p)$ such that $f\big( g(p), p\big) = 0$ for all $p$ i.e.~$x_S = g(p)$ is a steady state of $f$. The parameter transformation we want to generate is the function $p \mapsto \big(p, g(p)\big)$. Here, the set of outer parameters is the set of the reaction rates  $p$ whereas the set of inner parameters contains these rates and the corresponding steady states as initial value parameters. The root of $f$ must be determined numerically which is implemented based on \code{multiroot()} from the \pkg{rootSolve} package.

The following code is a reimplementation of the example above. The condition "open" is special in that the reflux rate is modified compared to the reflux rate by which the steady state is computed. This is solved by an event, flipping a switch variable from 0 to 1 and thereby replacing the rate \code{reflux} by \code{reflux_open}.

\begin{CodeChunk}
\begin{CodeInput}
# Add reactions
reactions <- NULL
reactions <- addReaction(reactions, "TCA_buffer", "TCA_cell",
		 rate = "import*TCA_buffer",
		 description = "Uptake")
reactions <- addReaction(reactions, "TCA_cell", "TCA_buffer",
		 rate = "export_sinus*TCA_cell",
		 description = "Sinusoidal export")
reactions <- addReaction(reactions, "TCA_cell", "TCA_cana",
		 rate = "export_cana*TCA_cell",
		 description = "Canalicular export")
reactions <- addReaction(reactions, "TCA_cana", "TCA_buffer",
		 rate = "(reflux*(1-switch) + reflux_open*switch)*TCA_cana",
		 description = "Reflux into the buffer")
reactions <- addReaction(reactions, "0", "switch",
		 rate = "0",
		 description = "Create a switch")

# Translate into ODE model
mymodel <- odemodel(reactions, modelname = "bamodel")
\end{CodeInput}
\end{CodeChunk}

For the implicit parameter transformation we need the ODE which is obtained by the \code{as.eqnvec()} command from the reactions, an \code{eqnlist} object. The Jacobian of $f$ is rank-deficient because the system has a conserved quantity $c$, i.e.~\code{TCA_tot = TCA_buffer + TCA_cana + TCA_cell}. Replacing one element of $f$ by $c - \verb!TCA_tot!$, the rank of the Jacobian is completed, the condition for the local existence of the implicit function $g(p)$ is satisfied and the steady state is parameterized by $p$ and the additional parameter \code{TCA_tot}.

\begin{CodeChunk}
\begin{CodeInput}
# Set up implicit parameter transformation
f <- as.eqnvec(reactions)[c("TCA_buffer", "TCA_cana", "TCA_cell")]
f["TCA_cell"] <- "TCA_buffer + TCA_cana + TCA_cell - TCA_tot"
pSS <- P(f, "TCA_tot", method = "implicit",
         compile = TRUE, modelname = "pfn")
\end{CodeInput}
\end{CodeChunk}

For the optimization, all outer parameters should still be log-parameters, an explicit parameter transformation. The final transformation will be a concatenation of the implicit and explicit transformations, \code{pSS*p}.

\begin{CodeChunk}
\begin{CodeInput}
# Set up explicit parameter transformation
innerpars <- unique(c(getParameters(mymodel),
                      getSymbols(observables),
                      getSymbols(f)))
trafo <- as.eqnvec(innerpars, names = innerpars)
trafo[reactions$states] <- "0"
trafo <- replaceSymbols(innerpars,
                        paste0("exp(", innerpars, ")"),
                        trafo)
p <- P(trafo)
\end{CodeInput}
\end{CodeChunk}

Both, the exchange of buffer and the opening of bile canaliculi by a Ca$^{2+}$/Mg$^{2+}$-free buffer are implemented as events. We therefore have different prediction functions for the two conditions.

\begin{CodeChunk}
\begin{CodeInput}
# Set up prediction function with events
event.buffer <- data.frame(var = "TCA_buffer", 
			   time = 0, 
			   value = 0, 
			   method = "replace")
event.open   <- data.frame(var = "switch", 
			   time = 0, 
			   value = 1, 
			   method = "replace")

x <- Xs(mymodel, 
	events = event.buffer, 
	condition = "standard") +
     Xs(mymodel, 
        events = rbind(event.buffer, event.open), 
	condition = "open")
\end{CodeInput}
\end{CodeChunk}

Although the observables have not changed compared to the set-up with purely explicit transformations, the observation function must be generated again because the new parameter \code{TCA_tot} has appeared. This parameter is not contained in the reactions. Therefore, we use the \code{states} and \code{parameters} arguments to explicitly inform the observation function about the states and parameters involved. Otherwise, parameter sensitivities are not propagated correctly.

\begin{CodeChunk}
\begin{CodeInput}
# Generate observation function with modified states/parameters
g <- Y(observables,
       states = reactions$states,
       parameters = setdiff(innerpars, reactions$states),
       compile = TRUE, modelname = "obsfn")
\end{CodeInput}
\end{CodeChunk}

Finally, the objective function is defined. The prediction function is now a concatenation of four functions.

\begin{CodeChunk}
\begin{CodeInput}
# Generate objective function
outerpars <- getParameters(p)
pouter <- structure(rep(-1, length(outerpars)), names = outerpars)
obj <- normL2(data, g*x*pSS*p) + constraintL2(pouter, sigma = 10)
\end{CodeInput}
\end{CodeChunk}

The same simulated data set has been fitted by the fully explit and the implicit/explicit model implementations. In both cases the global optimum is unique. Parameter profiles have been computed with both model implementations, shown in Figure~\ref{fig:allprofiles}.
\begin{figure}[ht]
	\centering
	\includegraphics[width = \textwidth]{images/figure8}
	\caption{Comparison of parameter profiles. The profile likelihood for the model without steady-state constraints, explicit steady-state constraints and implicit implementation of steady states is visualized by different colors.}
	\label{fig:allprofiles}
\end{figure}
The optima found by all three approaches are statistically compatible. The two implementations using the steady-state information show exactly the same profiles. Since one formulation is parameterized by \code{TCA_cell} whereas the other is parameterized by \code{TCA_tot}, the plot highlights one of the fundamental properties of the profile likelihood: invariance under reparameterization. In comparison to the profiles without steady-state information, the new profiles are narrower, meaning that the parameters are better identifiable. This was to be expected because we have reduced the dimension of the parameter space by one.

\end{document}